Generalized spaces for constructive algebra

### 2019-09-23 slides-herrsching2019.tex
Slides for Lecture III finished

>>>>

\section{Applications in constructive algebra}

\begin{frame}
  \centering\bigskip
  \includegraphics[height=8em]{calabi-yau}

  \large Lecture III:
  \hil{Applications in constructive algebra}

  \begin{hilblock}Without loss of generality, any reduced ring is a field.\end{hilblock}
\end{frame}

\begin{frame}{A remarkable sheaf}
  \justifying
  Let~$A$ be a reduced ring ($x^n = 0 \Rightarrow x = 0$).
  Then there is a certain sheaf~$A^\sim$ of rings on a certain locale~$X$ such
  that \ldots

  {\centering $A^\sim$ is close to~$A$:\par}
  \vspace*{-1.4em}
  \begin{varblock}{\textwidth}{}
    \justifying
    There is a \hil{canonical bijection}~$A \to A^\sim(X)$.

    $A^\sim$ inherits any property of~$A$ which is
    \hil{localization-stable}.

    A geometric sequent holds for~$A^\sim$ iff$^\star$ it holds for \hil{all
    stalks}~$A_{\mathfrak{f}}$.
  \end{varblock}

  {\centering $A^\sim$ has better properties than~$A$:\par}
  \vspace*{-1.6em}
  \setbeamercolor{block body}{bg=red!30}
  \setbeamercolor{structure}{fg=purple}
  \begin{varblock}{\textwidth}{}
    $A^\sim$ is a \hil{field}: $\forall x\?A^\sim\_ (\neg(\exists y\?A^\sim\_
    xy = 1) \Rightarrow x = 0)$.

    $A^\sim$ has \hil{$\boldsymbol{\neg\neg}$-stable equality}:
    $\forall x,y\?A^\sim\_ \neg\neg(x = y) \Rightarrow x = y$.

    \mbox{$A^\sim$ is \hil{anonymously Noetherian}.}\\[-1.2em]
  \end{varblock}

  {\centering This observation can be exploited to give \\ short, conceptual and
  constructive proofs.\par}
\end{frame}

\begin{frame}{Examples}
  \vspace*{-1.2em}
  \begin{columns}[t]
    \begin{column}[t]{0.58\textwidth}
      \centering

      \begin{varblock}{\textwidth}{Injective matrices}
        \justifying
        Let~$M$ be an injective matrix with more columns than rows over a ring~$A$.
        Then~$1 = 0$ in~$A$.
      \end{varblock}

      \justifying
      \textbf{Proof.} \bad{Assume not.} Then there is a \bad{maximal
      prime filter} $\fff \subseteq A$. The matrix is injective over the \bad{field}~$A_\fff$;
      contradiction to basic linear algebra.\medskip

      \textbf{Proof.} $M$ is also injective as a matrix over~$A^\sim$.
      This is a contradiction by basic linear
      algebra. Thus~$X \models \bot$. This amounts to~$1 = 0$ in~$A$.
    \end{column}

    \begin{column}[t]{0.46\textwidth}
      \centering

      \begin{varblock}{\textwidth}{Generic freeness\phantom{p}}
        \justifying
        Let~$M$ be a finitely generated~$A$-module.
        If~$f = 0$ is the only element of~$A$ such that~$M[f^{-1}]$ is a
        free~$A[f^{-1}]$-module, then~$1 = 0$ in~$A$.
      \end{varblock}

      \justifying
      \textbf{Proof.} See [Stacks Project].
      \medskip

      \textbf{Proof.} The claim amounts to \mbox{$X \models
      \text{``$M^\sim$}$}$\text{ is \hil{not not} free''}$. This statement
      follows from basic linear algebra over the field~$A^\sim$.
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[plain]
  \centering
  \includegraphics[height=\textheight]{lombardi-quitte}

  \jnote{1}{
    The book is \fixedhref{https://arxiv.org/abs/1605.04832}{available on the
    arXiv} and warmly recommended, as are all slides by Thierry Coquand on
    constructive algebra.«
  }
\end{frame}
\addtocounter{framenumber}{-1}
